---
title: '[Reproduced] Spurious Correlations and Random Walks'
author: yjtek
date: '2019-11-30'
slug: spurious-correlations-and-random-walks
categories: []
tags: []
draft: true
---

*Reproduced from [Fabian Dablander](https://fabiandablander.com/r/Spurious-Correlation.html)*
<br>
<br>

#### Introduction

This piece will explore the implications of random walk processes in inducing spurious correlations and false positives in hypothesis testing. Why can supposedly independent processes be heavily correlated?

Recall that an AR series generally takes the following form:  
<br>
<center> $Y_{t} = \phi Y_{t-1} + \epsilon_{t}$ --- (1) </center>
<br>
where $\phi$ measures the autocorrelation, and $\epsilon \sim \mathbb{N}(0,\sigma^{2})$. Note that a random walk is just a special case of the AR process where $\phi$ = 1.
<br>
<br>

#### Simulation

To explore the impact of $\phi$ in generating data, we build a function that simulates a data generating process given values of 

```{r}
simulate_ar <- function(n, phi, sigma = .1) {
  y <- rep(0, n)
  
  for (t in seq(2, n)) {
    y[t] <- phi*y[t-1] + rnorm(1, 0, sigma)
  }
  
  y
}
```

where $n$ determines the number of datapoints to generate; $\phi$ is the first order autocorrelation, and $\sigma$ is the variance of the errors in $\epsilon$.

As an example, let's generate 3 AR(1) processes and 3 random walks and chart them. Notice how much better behaved the AR processes are compared to the random walk.

```{r}
n <- 100
set.seed(1)
 
rw1 <- simulate_ar(n, phi = 1)
rw2 <- simulate_ar(n, phi = 1)
rw3 <- simulate_ar(n, phi = 1)

ar <- simulate_ar(n, phi = 0.5)
ar <- simulate_ar(n, phi = 0.5)
ar <- simulate_ar(n, phi = 0.5)
```



